Visual narratives of the Covid-19 pandemic

### 2022-01-24 JDSSV_manuscript_knitr.Rnw
updated section titles; shortened intro

A best practice for effective communication is to present information in an understandable way \citep{gigerenzerHelpingDoctorsPatients2007,gigerenzerWhatAreNatural2011}, for example by saying "one in ten" instead of 10\%. Using absolute rather than relative numbers, presenting information in an appealing graphic form and summarising the most important facts in "fact boxes" are some of the methods that have been developed and advocated to increase transparency in communication between stakeholders such as healthcare providers and patients (see for example: \url{https:// www.hardingcenter.de}).
Fact boxes in combination with icon arrays are recommended for the presentation of test results. Both representations are based on natural frequencies \citep{gigerenzerWhatAreNatural2011,Krauss2020} and represent case numbers as simply and concretely as possible. Many scientific studies show that icon arrays help people to understand numbers and risks more easily (e.g. \cite{McDowell2019}). The Harding Center for Risk Literacy shows many other examples of transparent communication of risks, including COVID-19\footnote{\url{https://www.hardingcenter.de/de/mrna-schutzimpfung-gegen-covid-19-fuer-aeltere-menschen}}.

Nevertheless, both the complexity of phenomena and the "bipolarity" of statistical thinking remain a challenge. While human thinking tends to simplify patterns and political communication also prefers a simple cause-effect relationship, real phenomena are often multivariate. Thus, in studying COVID-19 and predicting its spread, it is not only important to consider symptomatology, disease incidence and geographic distribution, population behaviour patterns, government policies, and impacts on the economy, schools, people in nursing homes, and society at large, but also to incorporate these into data analyses and communication of results. Associations observed in the data can often be caused by third party variables (confounders). In addition, much of the data comes from observational studies, which usually makes robust causal attribution problematic. However, statisticians who point out these limitations are at risk of having their statements pulled out one-sidedly in a polarised debate \citep{McConway2021}.

\section{The global perspective}

@@

Simulations were used particularly illustratively in the media in the course of the pandemic. An inspiring example illustrating the spread of the epidemic appeared as early as 14 March 2020 in the Washington Post\footnote{\url{https://www.washingtonpost.com/graphics/2020/world/corona-simulator/}} with the title "Why outbreaks like coronavirus spread exponentially, and how to flatten the curve". The Washington Post made this simulation available free of charge and in all major languages, which led to it being distributed worldwide, including repeatedly on German television\footnote{\url{https://web.br.de/interaktiv/corona-simulation/}}. The New York Times\footnote{\url{https://www.nytimes.com/interactive/2020/us/coronavirus-spread.html}} published a dynamic graphic entitled "How the Virus Won", which maps the spread of COVID-19 cases from February to June 2020 in the USA. It shows how an analysis of the associations between different COVID-19 strains and travel patterns can help understand the spread of the disease.

Another illustrative example is a simulation from ZEIT Online\footnote{\url{https://www.zeit.de/wissen/gesundheit/2020-11/coronavirus-aerosole-ansteckungsgefahr-infektion-hotspot-innenraeume}}, which - based on models developed by a group of researchers at the Max Planck Institute for Chemistry - estimates the probability of an infected person infecting other people in closed rooms in various scenarios. While the visualisations present the simulated infection processes in a catchy way, the dependence of the simulations on parameter assumptions and settings is usually not addressed. Simulations should also always make transparent on which model assumptions and which data basis the simulations were created.

\section{Time series}
% Line plots, frequency domain, NY times tile plots w/ cases represented in colour intensity

@@

\section{Comparisons and rankings}
\label{sec:rankings}

@@

>>>>

A best practice for effective communication is to present information in an understandable way \citep{gigerenzerHelpingDoctorsPatients2007,gigerenzerWhatAreNatural2011}, for example by saying "one in ten" instead of 10\%. Using absolute rather than relative numbers, presenting information in an appealing graphic form and summarising the most important facts in "fact boxes" are some of the methods that have been developed and advocated to increase transparency in communication between stakeholders such as healthcare providers and patients (see for example: \url{https:// www.hardingcenter.de}). Such "fact boxes" combined with the dynamic layer of the internet allowed for illustrative simulations of the spread of the epidemic, for example as presented already on 14 March 2020 in the Washington Post\footnote{\url{https://www.washingtonpost.com/graphics/2020/world/corona-simulator/}} with the title "Why outbreaks like coronavirus spread exponentially, and how to flatten the curve".  Nevertheless, both the complexity of phenomena and the "bipolarity" of statistical thinking remain a challenge. While human thinking tends to simplify patterns and political communication also prefers a simple cause-effect relationship, real phenomena are often multivariate. Thus, in studying COVID-19 and predicting its spread, it is not only important to consider symptomatology, disease incidence and geographic distribution, population behaviour patterns, government policies, and impacts on the economy, schools, people in nursing homes, and society at large, but also to incorporate these into data analyses and communication of results. Associations observed in the data can often be caused by third party variables (confounders). In addition, much of the data comes from observational studies, which usually makes robust causal attribution problematic. However, statisticians who point out these limitations are at risk of having their statements pulled out one-sidedly in a polarised debate \citep{McConway2021}.

\section{The global narrative}

@@

\section{The temporal narrative}
% Line plots, frequency domain, NY times tile plots w/ cases represented in colour intensity

@@

\section{The ranking narrative}
\label{sec:rankings}

@@

Ubiquity of the data, persisting general interest on the topic, the rise of data journalism, and the opportunities provided by interactive and dynamic internet platforms contributed to a an ongoing trend on visual narratives of the Covid-19 pandemic. Graphical displays are used as weapons for information, but also for misinformation. Despite a growing number of guidelines for communicating statistical results transportation of the intended messages is not always straightforward and remains a challenge. The deciphering of visual icons and best practices of visual communication require furhter studies on perception and graphics.
